State of the Practice of Lattice Boltzmann Method Software Development

### 2021-07-28 StateOfPractice/Peter-Notes/Report/Report.tex
updated LBM report, added rationale for recommendations

\noindent\textbf{NUMBER OF PAGES:} 121

@@

\begin{itemize}
	\item Include OS compatibility, including specific OS versions.
	\item Provide complete installation instructions.
	\item Installation instructions should be written as if the user does not have any dependencies installed and is installing on a clean OS. This can be tested on a clean environment using a virtual machine.
	\item List all dependencies in the installation instructions, dependency versions, and how to install them.
	\item If possible, automate the installation of dependencies. Use tools such as Make. 
	\item Automate the installation process as much as possible. Use tools such as Make.
	\item The installation instruction should only be in one location.
	\item Include descriptive error messages for errors encountered during installation. This was done by several top ranked packages, including Ludwig and LUMA
	\item Provide a way to validate the installation. This can be done using a custom script, or a test case that specifies expected output. This was done by several top ranked packages, including OpenLB, Ludwig, LUMA, and Palabos.
	\item Provide instructions for uninstalling the software. These were included with pyLBM.
\end{itemize}

@@

\begin{itemize}
	\item Use a requirements specification document. A potential template is presented in \citep{smith2005new}. 
	\item Make public (on package website or GitHub) the requirements specification document, or explicitly reference the domain theory that the software is designed from.
	\item Ensure the above information is easy to find. Consider adding it to the user manual. 
	\item Development teams should include both domain experts and experienced software developers.
	\item Use and make public (on package website or GitHub) detailed documentation. Consider using automatic document generation tools like Doxygen, Drasil, or Sphinx.
	\item Provide detailed tutorials that include expected output, like the \href{https://www.walberla.net/doxygen/index.html}{waLBerla tutorials}.
	\item Use unit tests during development and make them public (on package website or GitHub). This was available for Ludwig.
	\item Modularize the source code, separate components, hide information  behind well defined interfaces.
	\item Use continuous integration tools (Bamboo, Jenkins, and Travis CI) and processes during development. This was done by top ranked packages ESPResSo and Ludwig.
\end{itemize}

@@

\begin{itemize}
	\item Include descriptive error messages where appropriate. This was done by most packages that encountered a fault.
	\item In case automatic installation of dependencies fails, the system should indicate to the user what dependencies need to be installed manually. This was done by many packages, including ESPResSo++, Ludwig, LUMA, pyLBM, TCLB.
	\item The packages should include detailed tutorials, including dependencies, expected output, and any additional supplementary documentation that may be required. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, pyLBM.
\end{itemize}

@@

\begin{itemize}
	\item Include user hardware and software requirements documentation. Hardware requirements are rarely listed in the software packages that were assessed. Only those offering GPU processing (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB) mentioned any hardware requirements. On the other hand, some sort of software requirements were available for all packages, even if only some dependencies or a compatible operating system were mentioned.
	\item Provide a user tutorial that indicates the expected output. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, pyLBM.
	\item Provide a detailed user manual. It should identify elements of user interfaces, and identify all requirements to model a system. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item Provide documentation that details the background theory information, or provide a reference to such information.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and Palabos did this. The importance of specifying user characteristics is discussed in \citep{smith2007requirements}.
	\item Keep all documentation in one location.
	\item Do not duplicate artifact information.
	\item Maintain a user support model (Git, email, forum, FAQ)
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to shorten development time and improve usability.
\end{itemize}

@@

\begin{itemize}
	\item Keep artifacts updated.
	\item Ideally include most of the common and less common artifacts listed in Section \ref{artifacts}.
	\item Include version numbers and release notes for all major source code and artifact releases. The top five ranked software packages include release notes.
	\item Have a defined process for accepting contributions, and make public documentation for making contributions to the project. The top five ranked software packages include information on how to contribute. 
	\item Use an issue tracker (Git, email, SourceForge, other) to manage bugs and changes. Issues should be regularly reviewed and closed. All but three (laboetie, lettuce, Sailfish) of the software packages that use Git have most of their issues closed.
	\item Use a version control system (GitHib, CVS). Four (ESPREsSo, Ludwig, LUMA, Palabos) of the top five ranked packages use a version control system.
	\item Source code needs to be well commented. Typically, more than 10 percent of LBM package source code is comments, as presented in Table \ref{gitrepodata}. Four of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, OpenLB) have about 20 percent of their source code as comments.
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Eliminate code duplication.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including maintainability.
	\item Consider the recommended points for addressing traceability.
\end{itemize}

@@

\begin{itemize}
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including modifiability.
	\item Consider future source code modifiability as early as the design stage of development.
	\item Consider flexibility of data structures and data storage in the design stage. The package pyLBM redeveloped data structures to ease future changes.
\end{itemize}

@@

\begin{itemize}
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Document module interfaces in design and developer documentation. 
	\item Provide API documentation, if applicable. Only one (ESPResSo) of the top five ranked packages provided API documentation.
\end{itemize}

@@

\begin{itemize}
	\item Adopt a coding standard and document it in artifacts, including some examples. A coding standard needs to be part of continuous integration. Only
	LUMA and HemeLB explicitly identify coding standards. 
	\item Use consistent, distinctive, and meaningful code identifiers. Nineteen of the 23 software packages use such code identifiers.
	\item Use symbolic constants. Twelve of the 23 software packages use such constants.
	\item Identify algorithms that are used in source code comments. Document them in the artifacts. Cite relevant external sources if needed. Domain algorithms are noted in the source code of 11 packages.
	\item Add meaningful comments. Indicate what is being done in each section of source code.
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Provide a user manual that identifies all technical and problem modeling requirements. Thirteen software packages have a user manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide. 
	\item Consider adding a FAQ section to the documentation. This helped resolve some usability issues for OpenLB.
\end{itemize}

@@

\begin{itemize}
	\item Update all relevant documentation when a change to the software is made.
	\item Use automatic document generation tools (Doxygen, Drasil, Sphinx) to limit the time spent on updating documentation.
	\item Provide a developer's guide to help orient developers to the artifacts, source code, and system architecture so that they can better document changes.  
\end{itemize}

@@

\begin{itemize}
	\item Summarize the development process that is used. Provide information on how new users can contribute. Identify the development model by name (waterfall, agile, etc.), if appropriate.
	\item Identify the development environment. Eight of the 23 software packages identify the development environment.
	\item Update all relevant documentation when a change to the software is made.
	\item Include notes with all releases. Nine of the packages include release notes.
	\item Communicate all changes within the development team.
	\item Use continuous integration processes and tools (Bamboo, Jenkins, and Travis CI). This was done by top ranked packages ESPResSo and Ludwig.
	\item Consider peer review processes to assess contributions and ensure the tracking of information. High ranked package ESPResSo uses a peer review process for contributions. 
	\item Use project management tools, including change and version control tools (GitHub, GitLab, CVS), collaboration tools (GitHub, GitLab), and document generation tools (Doxygen, Drasil, Sphinx).
\end{itemize}

@@

\begin{itemize}
	\item Test output against automatically calculated or known correct results. High ranked package ESPResSo does this.
	\item Automate the testing of output. Run tests after all code changes. High ranked package ESPResSo does this.
	\item Consider peer review processes and task based inspection to assess contributions. High ranked package ESPResSo does this.
	\item Update all relevant design documentation when a change to the software is made.
\end{itemize}

@@

\begin{itemize}
	\item Documentation must state all technical requirements and software dependencies. A missing dependency was a frequent cause of fault conditions during this assessment.
	\item Provide a detailed user manual. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item Provide a table of symbols in the developer's guide that maps to names used in the course code.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item The documentation should either explain the underlying CFD theories or provide a reference to appropriate resources.
	\item Consider asking users for feedback on the documentation. The developer of pyLBM noted that such feedback would be appreciated.
\end{itemize}

@@

\begin{itemize}
	\item Provide a detailed user manual. It should identify elements of user interfaces, and identify all requirements to model a system.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. 
	\item Include detailed tutorials, including dependencies, expected
	output, and any additional supplementary documentation that may be required.
	\item Keep all documentation in one location.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. 
	\item Modularize the source code, separate components, hide information behind well defined interfaces.
	\item Include descriptive error messages where appropriate.
	\item Summarize the development process that is used. Provide information on how new users can contribute. 
	\item Consider peer review processes and task based inspection to assess contributions.	
	\item Use continuous integration tools (Bamboo, Jenkins, and Travis CI) and processes during development. 
	\item Use project management tools, including change and version control tools (GitHub, GitLab, CVS), collaboration tools (GitHub, GitLab), and document generation tools (Doxygen, Drasil, Sphinx).
\end{itemize}

>>>>

\noindent\textbf{NUMBER OF PAGES:} 123

@@

\begin{itemize}
	\item Include OS compatibility, including specific OS versions. This was done by several top ranked packages, including ESPResSo and LUMA.
	\item Provide complete installation instructions. This was done by all of the top five ranked packages.
	\item Installation instructions should be written as if the user does not have any dependencies installed and is installing on a clean OS. This was done by several top ranked packages, including ESPResSo, OpenLB, and Palabos. This can be tested on a clean environment using a virtual machine.
	\item List all dependencies in the installation instructions, dependency versions, and how to install them. This was done by several top ranked packages, including ESPResSo, LUMA, OpenLB, and Palabos.
	\item If possible, automate the installation of dependencies. Use tools such as Make. This was done by all of the top five ranked packages.
	\item Automate the installation process as much as possible. Use tools such as Make. This was done by all of the top five ranked packages.
	\item The installation instruction should only be in one location. This was done by several top ranked packages, including ESPResSo, OpenLB, and Palabos.
	\item Include descriptive error messages for errors encountered during installation. This was done by several top ranked packages, including Ludwig and LUMA
	\item Provide a way to validate the installation. This can be done using a custom script, or a test case that specifies expected output. This was done by several top ranked packages, including OpenLB, Ludwig, LUMA, and Palabos.
	\item Provide instructions for uninstalling the software. These were included with pyLBM.
\end{itemize}

@@

\begin{itemize}
	\item Use a requirements specification document. This is suggested in SCS literature, and several of the top ranked packages had such a document or reference to theory manuals. A potential template is presented in \citep{smith2005new}. 
	\item Make public (on package website or GitHub) the requirements specification document, or explicitly reference the domain theory that the software is designed from.  This was done by several top ranked packages, including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Ensure the above information is easy to find. Consider adding it to the user manual. The information is easy to find in most top ranked packages.
	\item Development teams should include both domain experts and experienced software developers. This suggestion is based on developer comments.
	\item Use and make public (on package website or GitHub) detailed documentation. Consider using automatic document generation tools like Doxygen, Drasil, or Sphinx. This was done by all of the top five ranked packages.
	\item Provide detailed tutorials that include expected output, like the \href{https://www.walberla.net/doxygen/index.html}{waLBerla tutorials}.
	\item Use unit tests during development and make them public (on package website or GitHub). This was available for Ludwig.
	\item Modularize the source code, separate components, hide information  behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Use continuous integration tools (Bamboo, Jenkins, and Travis CI) and processes during development. This was done by top ranked packages ESPResSo and Ludwig.
\end{itemize}

@@

\begin{itemize}
	\item Include descriptive error messages where appropriate. This was done by most packages that encountered a fault.
	\item In case automatic installation of dependencies fails, the system should indicate to the user what dependencies need to be installed manually. This was done by many packages, including ESPResSo++, Ludwig, LUMA, pyLBM, TCLB.
	\item The packages should include detailed tutorials, including dependencies, expected output, and any additional supplementary documentation that may be required. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, and pyLBM.
\end{itemize}

@@

\begin{itemize}
	\item Include user hardware and software requirements documentation. Hardware requirements are rarely listed in the software packages that were assessed. Only those offering GPU processing (ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB) mentioned any hardware requirements. On the other hand, some sort of software requirements were available for all packages, even if only some dependencies or a compatible operating system were mentioned.
	\item Provide a user tutorial that indicates the expected output. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, pyLBM.
	\item Provide a detailed user manual. It should identify elements of user interfaces, and identify all requirements to model a system. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item Provide documentation that details the background theory information, or provide a reference to such information. This was done by several top ranked packages, including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and Palabos did this. The importance of specifying user characteristics is discussed in \citep{smith2007requirements}.
	\item Keep all documentation in one location. This was done by top ranked packages.
	\item Do not duplicate artifact information. This is suggested as good software development practice. Duplicate information is difficult to maintain. 
	\item Maintain a user support model (Git, email, forum, FAQ). This was done by all of the top five ranked packages.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to shorten development time and improve usability.
\end{itemize}

@@

\begin{itemize}
	\item Keep artifacts updated. This is suggested as good software maintenance practice.
	\item Ideally include most of the common and less common artifacts listed in Section \ref{artifacts}.
	\item Include version numbers and release notes for all major source code and artifact releases. The top five ranked software packages include release notes.
	\item Have a defined process for accepting contributions, and make public documentation for making contributions to the project. The top five ranked software packages include information on how to contribute. 
	\item Use an issue tracker (Git, email, SourceForge, other) to manage bugs and changes. Issues should be regularly reviewed and closed. All but three (laboetie, lettuce, Sailfish) of the software packages that use Git have most of their issues closed.
	\item Use a version control system (GitHib, CVS). Four (ESPREsSo, Ludwig, LUMA, Palabos) of the top five ranked packages use a version control system.
	\item Source code needs to be well commented. Typically, more than 10 percent of LBM package source code is comments, as presented in Table \ref{gitrepodata}. Four of the top five overall ranked packages (Ludwig, ESPResSo, Palabos, OpenLB) have about 20 percent of their source code as comments.
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Eliminate code duplication. This is suggested as good software development practice. Duplicate code is difficult to maintain.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including maintainability.
	\item Consider the recommended points for addressing traceability.
\end{itemize}

@@

\begin{itemize}
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including modifiability.
	\item Consider future source code modifiability as early as the design stage of development. This is suggested as good software development practice.
	\item Consider flexibility of data structures and data storage in the design stage. The package pyLBM redeveloped data structures to ease future changes.
\end{itemize}

@@

\begin{itemize}
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Document module interfaces in design and developer documentation. This is suggested as good software development practice.
	\item Provide API documentation, if applicable. Only one (ESPResSo) of the top five ranked packages provided API documentation.
\end{itemize}

@@

\begin{itemize}
	\item Adopt a coding standard and document it in artifacts, including some examples. A coding standard needs to be part of continuous integration. Only
	LUMA and HemeLB explicitly identify coding standards. 
	\item Use consistent, distinctive, and meaningful code identifiers. Nineteen of the 23 software packages use such code identifiers.
	\item Use symbolic constants. Twelve of the 23 software packages use such constants.
	\item Identify algorithms that are used in source code comments. Document them in the artifacts. Cite relevant external sources if needed. Domain algorithms are noted in the source code of 11 packages.
	\item Add meaningful comments. Indicate what is being done in each section of source code. All of the top ranked packages have meaningful comments in their source code.
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Provide a user manual that identifies all technical and problem modeling requirements. Thirteen software packages have a user manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide. 
	\item Consider adding a FAQ section to the documentation. This helped resolve some usability issues for OpenLB.
\end{itemize}

@@

\begin{itemize}
	\item Update all relevant documentation when a change to the software is made. This is suggested as good software development practice.
	\item Use automatic document generation tools (Doxygen, Drasil, Sphinx) to limit the time spent on updating documentation. This was done by all of the top five ranked packages.
	\item Provide a developer's guide to help orient developers to the artifacts, source code, and system architecture so that they can better document changes. Top ranked package OpenLB has a developer guide. 
\end{itemize}

@@

\begin{itemize}
	\item Summarize the development process that is used. Provide information on how new users can contribute. Identify the development model by name (waterfall, agile, etc.), if appropriate.  Seven of the software packages have some artifacts outlining the general development process. Eleven packages have information on how to contribute. 
	\item Identify the development environment. Eight of the 23 software packages identify the development environment.
	\item Update all relevant documentation when a change to the software is made. This is suggested as good software development practice.
	\item Include notes with all releases. Nine of the packages include release notes.
	\item Communicate all changes within the development team. This suggestion is based on developer comments.
	\item Use continuous integration processes and tools (Bamboo, Jenkins, and Travis CI). This was done by top ranked packages ESPResSo and Ludwig.
	\item Consider peer review processes to assess contributions and ensure the tracking of information. High ranked package ESPResSo uses a peer review process for contributions. 
	\item Use project management tools, including change and version control tools (GitHub, GitLab, CVS), collaboration tools (GitHub, GitLab), and document generation tools (Doxygen, Drasil, Sphinx). This was done by all of the top five ranked packages. 
\end{itemize}

@@

\begin{itemize}
	\item Test output against automatically calculated or known correct results. High ranked package ESPResSo does this.
	\item Automate the testing of output. Run tests after all code changes. High ranked package ESPResSo does this.
	\item Consider peer review processes and task based inspection to assess contributions. High ranked package ESPResSo does this.
	\item Update all relevant design documentation when a change to the software is made. This is suggested as good software development practice.
\end{itemize}

@@

\begin{itemize}
	\item Documentation must state all technical requirements and software dependencies. A missing dependency was a frequent cause of fault conditions during this assessment.
	\item Provide a detailed user manual. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item Provide a table of symbols in the developer's guide that maps to names used in the source code. This is suggested as good software development practice.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item The documentation should either explain the underlying CFD theories or provide a reference to appropriate resources. This was done by several top ranked packages, including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Consider asking users for feedback on the documentation. The developer of pyLBM noted that such feedback would be appreciated.
\end{itemize}

@@

\begin{itemize}
	\item Provide a detailed user manual. It should identify elements of user interfaces, and identify all requirements to model a system. Thirteen packages have a user manual. ESPResSo has a well detailed manual.
	\item State appropriate fluid dynamics problems that the software is designed to model in the documentation, and explicitly state the limits of the software. This is done in the ESPResSo user guide.
	\item Include detailed tutorials, including dependencies, expected
	output, and any additional supplementary documentation that may be required. This was done by many packages, including waLBerla, Palabos, MechSys, LUMA, and pyLBM.
	\item Keep all documentation in one location. This was done by top ranked packages.
	\item If possible, consider using popular user-friendly software languages like Python. Especially consider this for parts of the source code that is likely to be modified or reviewed by users. ESPResSo and Sailfish use Python to address several software qualities, including maintainability, modifiability and usability.
	\item Modularize the source code, separate components, hide information behind well defined interfaces. This is suggested in SCS literature, and in developer comments.
	\item Include descriptive error messages where appropriate. This was done by most packages that encountered a fault.
	\item Summarize the development process that is used. Provide information on how new users can contribute. Seven of the software packages have some artifacts outlining the general development process. Eleven packages have information on how to contribute.
	\item Consider peer review processes and task based inspection to assess contributions. High ranked package ESPResSo does this.
	\item Use continuous integration tools (Bamboo, Jenkins, and Travis CI) and processes during development. This was done by top ranked packages ESPResSo and Ludwig.
	\item Use project management tools, including change and version control tools (GitHub, GitLab, CVS), collaboration tools (GitHub, GitLab), and document generation tools (Doxygen, Drasil, Sphinx). This was done by all of the top five ranked packages.
\end{itemize}
